Eliminating object prior-bias from sparse-projection tomographic reconstructions

### 2021-10-01 paper/changes.tex
Changes in Complexity Analysis, minor changes in changes.pdf

   We accept this point, and have now discussed the difficulties in
   using deep-learning techniques for the current problem in Sec.~7C
   of our paper. We also describe them here below.

\textcolor{red}{Comment from Sharat:  Since the material is in the text, there is no need to mention it here.}

  In a machine learning based setting, our goal of detecting new
  changes is a `prediction' problem in a continuous solution space
  i.e., \textit{``given intensities of a voxel at various time
    instants in a longitudinal setting and partial measurements of the
    voxel at the current time, what will be the intensity of the same
    voxel at the current time instant?''} This estimation can be
  learnt by a deep neural network if there are hundreds of labeled
  data.  However, generalization of this across multiple datasets
  poses a question mark.

  %% thousands of time instants at which the previous voxel
  %% intensities are known for each voxel. This amount of data is not
  %% available to us, and may be hard to acquire because each specimen
  %% is usually scanned only a couple of times in any longitudinal
  %% study.

  Further, a specific `event-of-interest' may occur just once in a
  longitudinal study and hence training on data of all previous time
  instants may be misleading. If however, we use measurements from
  identical and full-cycle longitudinal studies, including
  events-of-interest, of similar specimen (instead of the `same'
  specimen as we used), a deep-learning technique may be applied. We
  see this as an extension of using object-prior generated from the
  same object, and hence we have not explored this direction.

  Another avenue for using deep-networks \textit{within} our current
  work is to replace the currently used fixed eigenspace provided by
  PCA by a learnt feature basis provided by a network such as an
  autoencoder. For each dataset, an autoencoder may be trained to
  learn a specific set of latent features (this might again require
  atleast a few tens of volumes). We see this as a future direction of
  work.

@@

  We now present details of both time and space complexities of our technique in Section.7B of our paper.

Most of the computation time in our technique is used for the
generation of low quality eigen-space by reconstructing each of the
templates using TV and FDK. If the set of projection views for every
test is fixed a priori, the set of low quality reconstructions can be
produced offline and stored in order to save computational costs. With
this assumption, Table~\ref{table:compute_time} shows the computation
times for various stages of reconstruction of a 2D slice of Liver on
MATLAB 2019a on an AMD 2920X 12-Core Processor machine with 64GB
RAM. If however, new sets of projection views for every test are to be
allowed, the low quality reconstructions can still be performed
efficiently using parallelization since these reconstructions are
completely independent. Here below, we describe the time complexity of
spatially-varying prior technique.  If
\begin{itemize}
\item $N =$ number of voxels of the volume
\item $T =$ number of templates available, and
\item $F =$ time taken for iterative TV optimization,
\end{itemize}
then, the computational complexity is sum of the following
\begin{itemize}
\item $O(NT^2+T^3)$ for creating the eigenspace,
\item $O(NT^2)$ for deriving the weights map, and
\item time $F$  taken for iterative minimization
\end{itemize}
where $O(.)$ is the notation for `Big O' used in complexity analysis. 
In terms of space, the complexity is $O(T)$ in order to store the templates and their low quality reconstructions. We only need to store a single weights map volume regardless of the number of templates available.

\begin{table}[]
  \centering
  \caption{Computation times (in seconds) for various stages of reconstruction of a 2D slice of Liver on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. Here we assume that low quality eigen-spaces of the templates are computed offline.}
\begin{tabular}{|l|l|l|l|}
\hline
Backprojection             & TV                        & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map\end{tabular} & \begin{tabular}[c]{@{}l@{}}Final reconstruction\\ with spatially-varying \\ prior\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.40} & \multicolumn{1}{c|}{8.44} & \multicolumn{1}{c|}{0.04}                                             & \multicolumn{1}{c|}{46.73}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\end{itemize}

>>>>

   We accept this point, and have now discussed the difficulties in
   using deep-learning techniques for the current problem in Section.~7C
   of our paper.

@@

  We now present details of both time and space complexities of our technique in Section.~7B of our paper.

  \end{itemize}

### 2021-10-01 paper/post_tci.tex
Changes in Complexity Analysis, minor changes in changes.pdf

\begin{equation}
  \begin{split}
    J_{\boldsymbol{\alpha}}(\boldsymbol{\x}) \triangleq &\lVert\boldsymbol{\mathcal{R} x- y}\rVert_2^2  + \lambda_1TV(\boldsymbol{x}) \\
    &+\lambda_2\lVert(\boldsymbol{W}(\boldsymbol{x} - (\boldsymbol{\mu + V\alpha}))\rVert_2^2
  \end{split}
  \end{equation}

@@

\begin{equation}
\boldsymbol{\alpha} = [(\boldsymbol{WV})^T\boldsymbol{WV}]^{-1}\boldsymbol{V}^T(\boldsymbol{W}^T\boldsymbol{W}(\boldsymbol{x} - \boldsymbol{\mu}))
\end{equation}

@@

\section{Discussion}
\label{sec:discussion}
In this section, we discuss tuning of parameters, computational complexity of our method, and requirements for applying deep learning techniques in longitudinal few-view reconstruction.
\subsection{Tuning of parameters:}
$\lambda_1$, $\lambda_2$ and $k$ are the three
hyper-parameters that need to be chosen carefully for optimal
reconstruction. In our experiments, $\lambda_1$ was tuned to maximize
SSIM of the whole reconstructed test volume for TV reconstruction of
each dataset. This value was retained for spatially-varying
prior-based reconstruction as well. The value of $\lambda_2$ largely
depends on the amount of artefacts we aim to remove by using
object-prior at the cost of their dominance in the new regions. This
value was chosen to lie between $0-3$ for our datasets. Finally, the
hyper-parameter $k$ defines the sensitivity of the weights map to the
difference between test image and object-prior (projection of
test onto the space of object-prior). When $k=0$, our method 
behaves similar to the reconstruction using TV-only regularization. As $k$ increases, the weights map starts
capturing new changes in the test, at the cost of detecting a few
false positives i.e., false new changes. In other words, as the
weights-map becomes more sensitive to the difference between the test
and object-prior, it becomes more noisy. In order to visualize the
effect of the hyper-parameter $k$, we performed 2D reconstructions on
okra dataset for different values of
$k$. Fig.~\ref{fig:few_view_okra_2D_weights} shows the weights map
obtained for each of the $k$ values and
Fig.~\ref{fig:reconstructions_as_k_varies} shows the corresponding
final reconstructions. We run our experiments for a few values of $k$
and choose the best reconstruction based on our tolerance for noise
in the weights-map.
When a large number of templates are available, this parameter can be
chosen by assuming one of the templates as test and choosing the $k$
that gave the best reconstruction. Alternatively, in cases where one
wishes to completely avoid the use of this hyper-parameter, one can
construct a binary weights-map using a learning based method described
in the supplementary material~\cite{supp_paper}.

@@

\subsection{Computational Complexity:}

@@

by reconstructing each of the templates using TV and FDK. If the set of projection views for every test is fixed a priori, the set of low quality reconstructions can be produced offline and stored in order to save computational costs. With this assumption, Table~\ref{table:compute_time} shows the computation times for various stages of reconstruction of a 2D slice of Liver on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. If however, new sets of projection views for every test are to be allowed, the low quality reconstructions can still be performed efficiently using parallelization since these reconstructions are completely independent. Here below, we describe the
time complexity of spatially-varying prior technique.  If

@@

\item $N =$ number of voxels of the volume
\item $T =$ number of templates available, and
\item $F =$ time taken for iterative TV optimization,

@@

then, the computational complexity is sum of the following

@@

\item $O(NT^2+T^3)$ for creating the eigenspace,
\item $O(NT^2)$ for deriving the weights map, and

@@

In terms of space, the complexity is $O(T)$ in order to store the templates and their low quality reconstructions. We only need to store a single weights map volume regardless of the number of templates available.

\begin{table}[]
 \centering
  \caption{Computation times (in seconds) for various stages of reconstruction of a 2D slice of Liver on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. Here we assume that low quality eigen-spaces of the templates are computed offline.}
\begin{tabular}{|l|l|l|l|}
\hline
Backprojection             & TV                        & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map\end{tabular} & \begin{tabular}[c]{@{}l@{}}Final reconstruction\\ with spatially-varying \\ prior\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.40} & \multicolumn{1}{c|}{8.44} & \multicolumn{1}{c|}{0.04}                                             & \multicolumn{1}{c|}{46.73}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\subsection{Requirements for applying deep learning techniques}
Deep learning techniques are powerful tools to learn an unknown
parameteric function relating an input space and an output space. In such
settings, our current goal of detecting new changes is a `prediction'
problem in a continuous solution space i.e., \textit{``given
  intensities of a voxel at various time instants in a longitudinal
  setting and partial measurements of the voxel at the current time,
  what will be the intensity of the same voxel at the current time
  instant?''}. This estimation can be learnt by a deep neural network
if there are hundreds of labeled data.  However, generalization of
this across multiple datasets poses a question mark.

>>>>

\begin{equation}
  \begin{split}
    J_{\boldsymbol{\alpha}}(\boldsymbol{\x}) \triangleq &\lVert\boldsymbol{\mathcal{R} x- y}\rVert_2^2  + \lambda_1TV(\boldsymbol{x}) \\
    &+\lambda_2\lVert(\boldsymbol{W}(\boldsymbol{x} - (\boldsymbol{\mu + V\alpha}))\rVert_2^2
  \end{split}
    \label{eq:compute_x}
  \end{equation}

@@

\begin{equation}
  \boldsymbol{\alpha} = [(\boldsymbol{WV})^T\boldsymbol{WV}]^{-1}\boldsymbol{V}^T(\boldsymbol{W}^T\boldsymbol{W}(\boldsymbol{x} - \boldsymbol{\mu}))
  \label{eq:compute_alpha}
\end{equation}

@@

\section{Discussion}
\label{sec:discussion}
In this section, we discuss tuning of parameters, computational complexity of our method, and requirements for applying deep learning techniques in longitudinal few-view reconstruction.
\subsection{Tuning of parameters}
$\lambda_1$, $\lambda_2$ and $k$ are the three
hyper-parameters that need to be chosen carefully for optimal
reconstruction. In our experiments, $\lambda_1$ was tuned to maximize
SSIM of the whole reconstructed test volume for TV reconstruction of
each dataset. This value was retained for spatially-varying
prior-based reconstruction as well. The value of $\lambda_2$ largely
depends on the amount of artefacts we aim to remove by using
object-prior at the cost of their dominance in the new regions. This
value was chosen to lie between $0-3$ for our datasets. Finally, the
hyper-parameter $k$ defines the sensitivity of the weights map to the
difference between test image and object-prior (projection of
test onto the space of object-prior). When $k=0$, our method 
behaves similar to the reconstruction using TV-only regularization. As $k$ increases, the weights map starts
capturing new changes in the test, at the cost of detecting a few
false positives i.e., false new changes. In other words, as the
weights-map becomes more sensitive to the difference between the test
and object-prior, it becomes more noisy. In order to visualize the
effect of the hyper-parameter $k$, we performed 2D reconstructions on
okra dataset for different values of
$k$. Fig.~\ref{fig:few_view_okra_2D_weights} shows the weights map
obtained for each of the $k$ values and
Fig.~\ref{fig:reconstructions_as_k_varies} shows the corresponding
final reconstructions. We run our experiments for a few values of $k$
and choose the best reconstruction based on our tolerance for noise
in the weights-map.
When a large number of templates are available, this parameter can be
chosen by assuming one of the templates as test and choosing the $k$
that gave the best reconstruction. Alternatively, in cases where one
wishes to completely avoid the use of this hyper-parameter, one can
construct a binary weights-map using a learning based method described
in the supplementary material~\cite{supp_paper}.

@@

\subsection{Computational Complexity}

@@

by reconstructing each of the templates using TV and FDK. If the set of projection views for every test is fixed a priori, the set of low quality reconstructions can be produced offline and stored in order to save computational costs. With this assumption, Table~\ref{table:compute_time} shows the computation times for various stages of reconstruction of a 2D slice of Liver (refer code in~\cite{codeRepo}) on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. If however, new sets of projection views for every test are to be allowed, the low quality reconstructions can still be performed efficiently using parallelization since these reconstructions are completely independent. Here below, we describe the
time complexity of spatially-varying prior technique.  Let

@@

\item $N$ denote the number of voxels of test volume,
\item $L$ denote the number of templates available,
\item $M$ denote the number of pilot reconstruction algorithms used for the computation of weights map, and
\item $F$ denote the time taken for optimization.

@@

$F$ consists of time taken for alternate minimization of Eq.~\ref{eq:spatially_varying_prior}, which in turn consists of time taken for repeated cycles of a closed-form update of $\boldsymbol{\alpha}$ (Eq.~\ref{eq:compute_alpha}) and an iterative optimization of  $\boldsymbol{x}$ (Eq.~\ref{eq:compute_x}). Speed of iterative optimization is dependent on the number of imaging views as well. Therefore, the time computational complexity of our method is sum of the following

@@

\item $O(NL^2+L^3)$ for creating eigenspace,
\item $O(NL)*M$ for deriving weights map, and

@@

In terms of space, the complexity is $O(NL +N)$ to store each of the $L$ templates and their low quality reconstructions, and to store one weights map. We only need to store a single weights map volume regardless of the number of templates available.

\begin{table}[]
 \centering
  \caption{Computation times (in seconds) for various stages of reconstruction of a 2D slice of Liver (refer code in~\cite{codeRepo}) on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. Here we assume that low quality eigen-spaces of the templates are computed offline.}
\begin{tabular}{|l|l|l|l|}
\hline
Backprojection             & TV                        & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map\end{tabular} & \begin{tabular}[c]{@{}l@{}}Final reconstruction\\ with spatially-varying \\ prior\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.40} & \multicolumn{1}{c|}{8.44} & \multicolumn{1}{c|}{0.04}                                             & \multicolumn{1}{c|}{46.73}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\subsection{Requirements for applying deep learning techniques}
Deep learning techniques are powerful tools to learn an unknown
parameteric function relating an input space and an output space. In such
settings, our current goal of detecting new changes is a `prediction'
problem in a continuous solution space i.e., \textit{``given
  intensities of a voxel at various time instants in a longitudinal
  setting and partial measurements of the voxel at the current time,
  what will be the intensity of the same voxel at the current time
  instant?''}. This estimation can be learnt by a deep neural network
if there are hundreds of labeled data.  However, generalization of
this across multiple datasets poses a question mark.
